An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2022-02-01 latex_files/ch5/5-1-ma2.tex
added conclusion and the extensibility part

Second-order moving average (MA2) is a probabilistic model commonly
used for time series analysis. The observation at time $t$ is given
by:

@@

\noindent
The random variable $w_{k} \sim \mathcal{N}(0,1), k \in \mathbb{Z}$
represent white noise and $\theta_1, \theta_2$, the two parameters of
interest, the dependence from the previous observations. The number of
consecutive observations $T$ is a hyper-parameter of the model that in
our case is set at $T=100$. For guaranteeing that the inference
problem is identifiable, i.e.\ the likelihood has only one mode, we
use the prior proposed by \cite{Marin2012} given in the equation
\eqref{eq:ma2_prior}:

@@

\noindent
The observation vector $\yb_0 = (y_1, \ldots, y_{100})$ is generated
with $\thetab^*=(0.6, 0.2)$. The dimensionality of the output $\yb$ is
high, therefore we use summary statistics. Considering that the
ouptput vector represents a time-series signal, we choose as summary
statistics the autocovariances with $lag=1$ and $lag=2$, as shown in
equations \eqref{eq:ma2_summary_1} and \eqref{eq:ma2_summary_2}. The
final distance node is the squared euclidean
distance \eqref{eq:ma2_summary_4}.

@@

\subsubsection*{Perform the inference}

As in the previous example, we perform the inference using both
optimisation alternatives (gradient-based and Bayesian optimisation)
and, in the end, by extending the implementation with a Neural Network
surrogate model. In the end, we use the Rejection ABC algorithm for
evaluating our approaches; Rejection ABC is a robust method in terms
of accuracy and can be used for baseline comparison.

The implementation's extension is performed using a multi-layer
perceptron as local surrogate function, as illustrated in chapter
... This intervention can be done easily by coding a custom
optimisation function with the surrogate model of our own
preference. The surrogate model substitutes the real distance function
$d_i$ inside the proposal region $q_i \quad \forall i$ at the
inference phase i.e. sampling, computing an expectation and evaluating
the posterior. In our example we use a neural network of two hidden
layers of 10 neurons each and we train it in 500 examples for each
proposal region.

In figure \ref{fig:ma2_5}, we illustrate the acceptance region of the
same deterministic simulator, in the gradient-based and the Bayesian
optimisation case. The acceptance regions are similar even though the
optimal points differ significantly due to the different optimisation
schemens.

In figure \ref{fig:ma2_3}, we demonstrate the histograms of the
marginal posteriors, using all four approaches; (a) Rejection ABC
(first column), (b) ROMC with gradient-based optimisation (second
column) (c) ROMC with Bayesian optimisation (third column) and (d)
ROMC with the Neural Network extendion. We observe that there is a
significant aggreement between the different approaches. The Rejection
ABC inference has been set to infer 10000 accepted samples, with
threshold $\epsilon=0.1$. At table \ref{tab:ma2} we present the
empirical mean $\mu$ and standard deviation $\sigma$ for each
inference approach. In figure \ref{fig:ma2_4}, we illustrate the
unnormalized posterior for the three different variations of the ROMC
method.

\subsubsection*{Extending the implementation}

\begin{center} \label{tab:ma2}
\begin{tabular}{ c|c|c|c|c }
\hline
& $\mu_{\theta_1}$ & $\sigma_{\theta_1}$ & $\mu_{\theta_2}$ & $\sigma_{\theta_2}$ \\
\hline \hline
Rejection ABC & 0.516 & 0.142 & 0.07 & 0.172 \\
\hline
ROMC (gradient-based) & 0.503 & 0.143 & 0.032 & 0.17 \\
\hline
ROMC (Bayesian optimisation) & 0.494 & 0.16 & 0.086 & 0.167 \\
\hline
ROMC (Neural Network) & 0.503 & 0.140 & 0.03 & 0.171 \\
\hline
\end{tabular}
\end{center}

% \begin{figure}[H]
%     \begin{center}
%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_distance_hist.png}
%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_distance_hist_bo.png} \\
%     \end{center}
%     \caption[MA2 example, histogram of distances]{Histogram of distances
%       $d_i^*, i \in \{ 1, \ldots, n_1 \}$. The left graph corresponds
%       to the gradient-based approach and the right one to the Bayesian
%       optimisation approach.}
%   \label{fig:ma2_2}
% \end{figure}

@@

\begin{figure}[ht]
  \begin{center}
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t1_rejection.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t1_romc.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t1_romc_nn.tex}
    }\\
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t2_rejection.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t2_romc.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t2_romc_nn.tex}
    }
    \end{center}
    \caption[MA2 example, evaluation of the marginal
    distributions.]{Histogram of the marginal posterior distributions
      using three different inference approaches; (a) in the first
      row, the samples are obtained using Rejection ABC sampling (b)
      in the second row, using ROMC with a gradient-based optimiser
      and (c) in the third row, using ROMC with Bayesian optimization
      approach. The vertical (red) line represents the samples mean
      $\mu$ and the horizontal (black) the standard deviation
      $\sigma$.}
  \label{fig:ma2_3}
\end{figure}

@@

>>>>

\subsubsection*{Model Definition}

MA2 is a probabilistic model for time series analysis. The observation
at time \(t\) is given by,

@@

\noindent
The r.v. \(w_{k} \sim \mathcal{N}(0,1) \) is white noise and the two
parameters of interest, \(\theta_1, \theta_2\), model the dependence
from the previous observations. The number of sequential observations
\(T\) is a hyper-parameter, in our case set to \(T=100\). For securing
that the inference problem is identifiable, i.e.\ the likelihood has
only one mode, we use the prior proposed by~\cite{Marin2012},

@@

\noindent
The observation vector \(\yb_0 = (y_1, \ldots, y_{100})\) is generated
with \(\thetab^*=(0.6, 0.2)\). The dimensionality of the output
\(\yb\) is high, therefore we use summary statistics. Considering that
the ouptput vector represents a time-series signal, we select the
autocovariances with \(\mathrm{lag}=1\) and \(\mathrm{lag}=2\), as shown in equations
\eqref{eq:ma2_summary_1} and \eqref{eq:ma2_summary_2}. The final
distance node is the squared euclidean distance
\eqref{eq:ma2_summary_4}.

@@

\subsubsection*{Inference}

In order to show all the capabilities of the implementation, we
perform the inference (i) using the gradient-based optimizer, (ii)
using the Bayesian Optimization scheme and (iii) fitting a Neural
Network (NN) as a surrogate model. We use the Rejection ABC algorithm
for evaluating all approaches. Replacing the typical quadratic
surrogate model with a NN serves as an illustrator of the
extensibility of our implementation. The replacement is done by coding
a custom optimisation function with a surrogate model of our own
preference, as shown in Chapter~\ref{subsec:extensibility}. For the
definition of the NN, we use the \pkg{MLPRegressor} class of the
\pkg{scikit-learn} package. Therefore, the NN substitutes the real
distance function \(d_i\) inside the proposal region
\(q_i \quad \forall i\) at the inference phase i.e. sampling,
computing an expectation and evaluating the posterior. In our example
we use a neural network of two hidden layers of 10 neurons each and we
train it sampling 500 examples from each proposal region.

In figure \ref{fig:ma2_5}, we illustrate the acceptance region of the
same deterministic simulator, in the gradient-based and the Bayesian
optimisation case. The acceptance regions are quite similar even though the
different optimisation schemes lead to different optimal points.

In figure \ref{fig:ma2_3}, we demonstrate the histograms of the
marginal posteriors, for each approach; (a) Rejection ABC (first
column), (b) ROMC with gradient-based optimisation (second column) (c)
ROMC with Bayesian optimisation (third column) and (d) ROMC with the
NN extension. We observe a significant aggreement between the
different approaches. At table \ref{tab:ma2} we present the empirical
mean \(\mu\) and standard deviation \(\sigma\) for each inference
approach and finally, in figure \ref{fig:ma2_4}, we illustrate the
unnormalized posterior for the three different variations of the ROMC
method. The results show that all ROMC variations provide consistent
results between them and in comparison with the standard Rejection ABC
algorithm.

\begin{center} \label{tab:ma2}
\begin{tabular}{ c|c|c|c|c }
\hline
& \(\mu_{\theta_1}\) & \(\sigma_{\theta_1}\) & \(\mu_{\theta_2}\) & \(\sigma_{\theta_2}\) \\
\hline \hline
Rejection ABC & 0.516 & 0.142 & 0.07 & 0.172 \\
\hline
ROMC (gradient-based) & 0.503 & 0.143 & 0.032 & 0.17 \\
\hline
ROMC (Bayesian optimisation) & 0.494 & 0.16 & 0.086 & 0.167 \\
\hline
ROMC (Neural Network) & 0.503 & 0.140 & 0.03 & 0.171 \\
\hline
\end{tabular}
\end{center}

% \begin{figure}[H]
%     \begin{center}
%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_distance_hist.png}
%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_distance_hist_bo.png} \\
%     \end{center}
%     \caption[MA2 example, histogram of distances]{Histogram of distances
%       \(d_i^*, i \in \{ 1, \ldots, n_1 \}\). The left graph corresponds
%       to the gradient-based approach and the right one to the Bayesian
%       optimisation approach.}
%   \label{fig:ma2_2}
% \end{figure}

@@

\begin{figure}[ht]
  \begin{center}
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t1_rejection.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t1_romc.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t1_romc_nn.tex}
    }\\
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t2_rejection.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t2_romc.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t2_romc_nn.tex}
    }
    \end{center}
    \caption[MA2 example, evaluation of the marginal
    distributions.]{Histogram of the marginal posterior distributions
      using three different inference approaches; (a) in the first
      row, the samples are obtained using Rejection ABC sampling (b)
      in the second row, using ROMC with a gradient-based optimiser
      and (c) in the third row, using ROMC with Bayesian optimization
      approach. The vertical (red) line represents the samples mean
      \(\mu\) and the horizontal (black) the standard deviation
      \(\sigma\).}
  \label{fig:ma2_3}
\end{figure}

@@

\subsubsection*{Parallelization}

As already stated, it is straightforward for the two basic parts of
the training part to be parallelised. There is a batch of \(n_1\)
independent objectives functions to be optimised and \(n_1\) bounding
boxes to be built. Both tasks can be applied in a parallel utilizing
all the available CPU cores. Our implementation supports
parallelization using the built-in \proglang{Python} package
\pkg{multiprocess}. In figure~\ref{fig:exec_parallel} we observe the
execution times for performing the inference. The parallel version
performs all tasks between 2.5 and 6 times faster compared to the
sequential; the optimization problems are solved almost 6 times
faster, sampling 3.5 times faster and evaluating the posterior and
constructing the bounding boxes almost 2.5 times faster.

\begin{figure}[ht]
  \centering
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Solve optimization problems},
          xlabel={\(n_1\)},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Proposal regions},
          xlabel={\(n_1\)},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Sample},
          xlabel={\(n_1\)},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Evaluate posterior},
          xlabel={\(n_1\)},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \caption[Execution time exploiting parallelization]{In the first
      line, we compare the parallel and sequential execution of the
      training part and at the second line the inference part. At the
      left figure, we measure the execution time for sampling \(n_2=50\)
      points per region. At the right figure we measure the execution
      time for evaluating the posterior at a batch of \(50\) points.}
      \label{fig:exec_parallel}
\end{figure}

### 2022-02-01 latex_files/ch5/5-2-execution-time.tex
added conclusion and the extensibility part

As already stated, it is straightforward for the two basic parts of
the training part to be parallelised. There is a batch of $n_1$
independent objectives functions to be optimised and $n_1$ bounding
boxes to be built. Both tasks can be applied in a parallel utilizing
all the available CPU cores. Our implementation supports
parallelization using the built-in \proglang{Python} package
\pkg{multiprocess}. In figure~\ref{fig:exec_parallel} we observe the
execution times for performing the inference. The parallel version
performs all tasks between 2.5 and 6 times faster compared to the
sequential. Optimisation problems are solved almost 6 times
faster. Sampling is executed 3.5 times faster, whereas evaluating the
posterior and constructing the bounding boxes almost 2.5 times faster.

\begin{figure}[ht]
  \centering
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Solve optimization problems},
          xlabel={$n_1$},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Proposal regions},
          xlabel={$n_1$},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Sample},
          xlabel={$n_1$},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Evaluate posterior},
          xlabel={$n_1$},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \caption[Execution time exploiting parallelization]{In the first
      line, we compare the parallel and sequential execution of the
      training part and at the second line the inference part. At the
      left figure, we measure the execution time for sampling $n_2=50$
      points per region. At the right figure we measure the execution
      time for evaluating the posterior at a batch of $50$ points.}
      \label{fig:exec_parallel}
\end{figure}

>>>>

As already stated, it is straightforward for the two basic parts of
the training part to be parallelised. There is a batch of \(n_1\)
independent objectives functions to be optimised and \(n_1\) bounding
boxes to be built. Both tasks can be applied in a parallel utilizing
all the available CPU cores. Our implementation supports
parallelization using the built-in \proglang{Python} package
\pkg{multiprocess}. In figure~\ref{fig:exec_parallel} we observe the
execution times for performing the inference. The parallel version
performs all tasks between 2.5 and 6 times faster compared to the
sequential. Optimisation problems are solved almost 6 times
faster. Sampling is executed 3.5 times faster, whereas evaluating the
posterior and constructing the bounding boxes almost 2.5 times faster.

\begin{figure}[ht]
  \centering
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Solve optimization problems},
          xlabel={\(n_1\)},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Proposal regions},
          xlabel={\(n_1\)},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Sample},
          xlabel={\(n_1\)},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \begin{minipage}[t]{.49\textwidth}
      \begin{tikzpicture}
        \begin{axis}[
          title={Evaluate posterior},
          xlabel={\(n_1\)},
          ylabel={seconds},
          xmin=1, xmax=500,
          ymin=0, ymax=7,
          xtick={1,100,200,300,400,500},
          ytick={0, 1, 2, 3, 4, 5, 6, 7},
          legend pos=north west,
          ymajorgrids=true,
          grid style=dashed,
          ]
          \addplot[color=blue, mark=o] coordinates {
            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
          };
          \addlegendentry{sequential}
          \addplot[color=red, mark=o] coordinates {
            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
          };
          \addlegendentry{parallel}          
        \end{axis}
      \end{tikzpicture}
    \end{minipage}\hfill
    \caption[Execution time exploiting parallelization]{In the first
      line, we compare the parallel and sequential execution of the
      training part and at the second line the inference part. At the
      left figure, we measure the execution time for sampling \(n_2=50\)
      points per region. At the right figure we measure the execution
      time for evaluating the posterior at a batch of \(50\) points.}
      \label{fig:exec_parallel}
\end{figure}

### 2022-02-01 paper.tex
added conclusion and the extensibility part

In this section, we illustrate the implementation performing two
experiments; In chapter~\ref{subsec:ma2} we perform the inference at
the second-order moving average (MA2) model, which is a standard model
of \pkg{ELFI}, and in chapter~\ref{subsec:exec_time} we measure the
execution speed-up achieved by the parallelization of ROMC.

\subsection{Moving average example}
\label{subsec:ma2}
\input{latex_files/ch5/5-1-ma2.tex}

\subsection{Execution time}
\label{subsec:exec_time}
\input{latex_files/ch5/5-2-execution-time.tex}

@@

>>>>

In this section, we test the implementation using the second-order
moving average (MA2) example, which is a standard model of
\pkg{ELFI}. We perform the inference using three different versions of
ROMC; (i) using a gradient-based optimizer, (ii) using the Bayesian
Optimization scheme and (iii) fitting a Neural Network as a surrogate
model. The later illustrates how to extend the implementation,
replacing part of ROMC with a user-defined component. Finally, we
measure the execution speed-up achieved by the parallelized version of
ROMC.

% \subsection{Moving average example}
% \label{subsec:ma2}
\input{latex_files/ch5/5-1-ma2.tex}

% \subsection{Execution time}
% \label{subsec:exec_time}
% \input{latex_files/ch5/5-2-execution-time.tex}

@@

In this paper, we presented the implementation details we followed for
developing the LFI method ROMC at the \proglang{ELFI} package. We paid
thorough attention to two specific use-case scenarios. Firstly, we
illustrate how a user may take advantage of our ready-to-use API for
solving its LFI problem. Secondly, we focus on the scenario where a
researcher wants to intervene and alter parts of the method. Our
implementation is designed to support this as well.

By all means, there are still open challenges for the future. Two
directions may enable ROMC to solve high-dimensional problems
efficiently. The first one is enabling ROMC's execution into a cluster
of computers. ROMC can be characterized as an \textit{embarrassingly
  parallel} workload; each optimization problem is an entirely
independent task. Therefore, supporting inference into a cluster of
computers can radically speed up the inference. The second one refers
to implementing the method in a framework that supports automatic
differentiation. Automatic differentiation is necessary for
efficiently solving optimization problems, especially in
high-dimensional parametric models.
